\chapter*{Acknowledgements}
\label{cha:section_acknowledgements}
To begin, I would like to acknowledge the support of the excellent Department of Physics and Astronomy at UNC Chapel Hill, for its vibrant and closely-knit community of students and faculty, its many excellent and brilliant professors, and the wonderful atmosphere of learning and collaboration and that is facilitated by everyone within. Without such a positive and encouraging context to work in, my research would've never become what it is today. Similarly, I would also like to thank current and past members of the Reichart Lab/Skynet Robotic Telescope Network for their support and collaboration throughout my time in the Lab, including Dylan Dutton, Vladimir Kouprianov, John Martin, and Roark Habeggar. I would also like to specifically thank Prof. Fabian Heitsch, as both his course on physical modeling, and my consulting of his vast knowledge of numerical techniques and modeling for this work equpped me tackle a number of challenging problems. I would also like to acknowledge the financial support of the NASA/North Carolina Space Grant Undergraduate Research Fellowship and Scholarship Programs, as well as the UNC Department of Physics and Astronomy in the form of the Earl Nelson Mitchell Scholarship in Physics, and the Robert Shelton Award for Outstanding Research.


% My friends/colleagues in physics... Thomases, Charlie, etc (maybe not add this incase of exclusion)?

Without Adam Trotter, this thesis would not exist, and not just because his 2011 PhD dissertation serves as the foundation for this work. Ever since I began work on this project, Adam has been incredibly helpful and insightful. His attentiveness and rapid response time towards my constant volley of questions is only matched by and his immediate dedication towards solving any problem I level at him. I very much look forward to continuing to work with Adam in the future.

Dr. Daniel Reichart. Prof. Reichart, or Dan, is a man of many names and of many interests. A theorist by training, he has worked with observational and radio astronomy, statistics, astronomy teaching, and many other disciplines. I am continually amazed at the veritable font of knowledge that he possesses, and his willingness to dive deep into any problem, straight to the whiteboard, no matter how difficult it becomes. I have never met someone with such rapid, full attentiveness and masterful way of elucidating difficult topics and quickly figuring out (often unorthodox) solutions to new problems. Overall, Dan has not just been indispensable to helping me with this project; he has been the driving motivator for my entire research career over these past three years. Anybody would be extremely lucky to have an advisor half as good as him.
% \glsresetall %% all glossary entries should be used in long form (again)
%% vim:foldmethod=expr
%% vim:fde=getline(v\:lnum)=~'^%%%%\ .\\+'?'>1'\:'='
%%% Local Variables:
%%% mode: latex
%%% mode: auto-fill
%%% mode: flyspell
%%% eval: (ispell-change-dictionary "en_US")
%%% TeX-master: "main"
%%% End:
